---
title: "Analysis - Change Over Time"
output: html_notebook
---

Loading in NC county-level Table B25048 from the 2018 5-year ACS (downloaded from data.census.gov). 
```{r}
setwd("/Users/korie/Documents/Fall 2020/MEJO390/final-project-korie/")
county_plumbing_2018 <- "data/ACSDT5Y2018.B25048_2020-10-11T194219 2/ACSDT5Y2018.B25048_data_with_overlays_2020-10-11T194213.csv"
library(tidyverse)
county_plumbing_2018 <- read_csv(county_plumbing_2018, skip=1)
colnames(county_plumbing_2018)
names(county_plumbing_2018)[names(county_plumbing_2018) == "Estimate!!Total"] <- "total_occupied_homes_2018"
names(county_plumbing_2018)[names(county_plumbing_2018) == "Margin of Error!!Total"] <- "moe_total_occupied_homes_2018"
names(county_plumbing_2018)[names(county_plumbing_2018) == "Estimate!!Total!!Complete plumbing facilities"] <- "est_complete_plumbing_2018"
names(county_plumbing_2018)[names(county_plumbing_2018) == "Margin of Error!!Total!!Complete plumbing facilities"] <- "moe_complete_plumbing_estimate_2018"
names(county_plumbing_2018)[names(county_plumbing_2018) == "Estimate!!Total!!Lacking complete plumbing facilities"] <- "est_lacking_complete_plumbing_2018"
names(county_plumbing_2018)[names(county_plumbing_2018) == "Margin of Error!!Total!!Lacking complete plumbing facilities"] <- "moe_lacking_complete_plumbing_2018" 
```

Loading in NC county-level Table B25048 from the 2013 5-year ACS (downloaded from data.census.gov) and cleaning column names. 
```{r}
setwd("/Users/korie/Documents/Fall 2020/MEJO390/final-project-korie/")
county_plumbing_2013 <- "data/ACSDT5Y2013.B25048_2020-10-11T194043/ACSDT5Y2013.B25048_data_with_overlays_2020-10-11T194011.csv"
county_plumbing_2013 <- read_csv(county_plumbing_2013, skip=1)
colnames(county_plumbing_2013)
names(county_plumbing_2013)[names(county_plumbing_2013) == "Estimate!!Total"] <- "total_occupied_homes_2013"
names(county_plumbing_2013)[names(county_plumbing_2013) == "Margin of Error!!Total"] <- "moe_total_occupied_homes_2013"
names(county_plumbing_2013)[names(county_plumbing_2013) == "Estimate!!Total!!Complete plumbing facilities"] <- "est_complete_plumbing_2013"
names(county_plumbing_2013)[names(county_plumbing_2013) == "Margin of Error!!Total!!Complete plumbing facilities"] <- "moe_complete_plumbing_estimate_2013"
names(county_plumbing_2013)[names(county_plumbing_2013) == "Estimate!!Total!!Lacking complete plumbing facilities"] <- "est_lacking_complete_plumbing_2013"
names(county_plumbing_2013)[names(county_plumbing_2013) == "Margin of Error!!Total!!Lacking complete plumbing facilities"] <- "moe_lacking_complete_plumbing_2013"
```

Joining the 2018 and 2013 county-level data in order to allow for change over time and percent change formulas to be used.
```{r}
library(dplyr)
plumbing_2018_2013 <- left_join(county_plumbing_2018, county_plumbing_2013, by="Geographic Area Name")
```

Calculating raw change and percent change in number of homes without adequate plumbing from 2013 to 2018.
```{r}
change <- plumbing_2018_2013 %>%
  mutate(raw_lacking_change = (est_lacking_complete_plumbing_2018 - est_lacking_complete_plumbing_2013),
         pct_lacking_change = ((est_lacking_complete_plumbing_2018 - est_lacking_complete_plumbing_2013)/(est_lacking_complete_plumbing_2018))*100) %>%
  arrange(desc(pct_lacking_change)) %>%
  select(`Geographic Area Name`, est_lacking_complete_plumbing_2018, moe_lacking_complete_plumbing_2018,  est_lacking_complete_plumbing_2013, moe_lacking_complete_plumbing_2013, raw_lacking_change, pct_lacking_change)

change
```
Number of homes in NC without proper plumbing decreased from 2013 to 2018 by 3,072, or 24.49%. 38 counties had the estimated number of homes go up from 2013, one county (Alamance) stayed the same, and 61 counties went down.











