---
title: "Visualizations"
output: html_notebook
---

Loading libraries.
```{r}
library(tidycensus)
library(censusapi)
```

Loading Census API key.
```{r}
Sys.setenv(CENSUS_KEY="f8926d2beb514da70fdcd1abdac3eab3c261b34a")
readRenviron("~/.Renviron")
Sys.getenv("CENSUS_KEY")
```

Pulling NC plumbing data directly from Census.
```{r}
nc_plumbing_county_2018 <- getCensus(name="acs/acs5", vintage=2018, vars=c("NAME", "B25048_003E"), region="county:*", regionin="state:37")
nc_plumbing_county_2013 <- getCensus(name="acs/acs5", vintage=2013, vars=c("NAME", "B25048_003E"), region="county:*", regionin="state:37")
nc_plumbing_tract_2018 <- getCensus(name="acs/acs5", vintage=2018, vars=c("NAME", "B25048_003E"), region="tract:*", regionin="state:37")
nc_plumbing_tract_2013 <- getCensus(name="acs/acs5", vintage=2018, vars=c("NAME", "B25048_003E"), region="tract:*", regionin="state:37")
```

Joining 2018 county data with NC shapefile in order to map.
```{r}
library(tigris)
options(tigris_class="sf")
nc_counties <- counties("NC", cb=T)
viz_county_nc_plumbing_2018 <- left_join(nc_counties, nc_plumbing_county_2018, by=c("COUNTYFP" = "county")) 
```

Mapping inadequate plumbing by county for 2018.
```{r}
library(ggplot2)
ggplot(viz_county_nc_plumbing_2018) +
  geom_sf(aes(fill=B25048_003E), color="white") +
  theme_void() +
  theme(panel.grid.major=element_line(colour="transparent")) +
  scale_fill_distiller(palette="Blues", direction=1, name="Number of homes") +
  ggtitle("Homes lacking complete plumbing facilities in N.C.") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(caption = "Source: U.S. Census/ACS5 2018")
```

Joining 2018 tract data with NC shapefile in order to map.
```{r}
library(tigris)
options(tigris_class="sf")
nc_tracts <- tracts("NC", cb=T)
viz_nc_tract_plumbing_2018 <- left_join(nc_tracts, nc_plumbing_tract_2018, by=c("TRACTCE" = "tract")) 
```

Mapping inadequate plumbing by tract for 2018.
```{r}
library(ggplot2)
ggplot(viz_nc_tract_plumbing_2018) +
  geom_sf(aes(fill=B25048_003E), color="white") +
  theme_void() +
  theme(panel.grid.major=element_line(colour="transparent")) +
  scale_fill_distiller(palette="Blues", direction=1, name="Number of homes") +
  ggtitle("Homes lacking complete plumbing facilities in N.C.") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  labs(caption = "Source: U.S. Census/ACS5 2018")
```

Joining 2018 and 2013 county files to map change over time.
```{r}
county_plumbing_change <- left_join(nc_plumbing_county_2018, nc_plumbing_county_2013, by=c("county" = "county"))
names(county_plumbing_change)[names(county_plumbing_change) == "B25048_003E.x"] <- "est_inadequate_plumbing_2018"
names(county_plumbing_change)[names(county_plumbing_change) == "B25048_003E.y"] <- "est_inadequate_plumbing_2013"
names(county_plumbing_change)[names(county_plumbing_change) == "state.x"] <- "state"
names(county_plumbing_change)[names(county_plumbing_change) == "NAME.x"] <- "county_name"
county_plumbing_change <- county_plumbing_change %>%
  select("state", "county", "county_name", "est_inadequate_plumbing_2018", "est_inadequate_plumbing_2013") %>%
  mutate(change = est_inadequate_plumbing_2018 - est_inadequate_plumbing_2013,
         change_category = case_when(change >0 ~ "increase",
                                    change <0 ~ "decrease",
                                    TRUE ~ "same"))
```

Joining county_plumbing_change with NC shapefile to create visualization.
```{r}
plumbing_change_with_shape <- left_join(nc, county_plumbing_change, by=c("COUNTYFP"="county"))
```

Mapping. 
```{r}
library(viridis)
library(ggplot2)
ggplot(plumbing_change_with_shape) +
  geom_sf(aes(fill=change_category), color="white") +
    theme_void() +
   theme(panel.grid.major=element_line(colour="transparent")) +
  ggtitle("Change in number of homes with inadequate plumbing, 2013 to 2018") +
   theme(plot.title = element_text(hjust = 1)) +
  labs(caption = "Source: U.S. Census/ACS5 2018, ACS5 2013") 
```

Creating a map with inadequate plumbing by county with a rural/urban overlay.



